---
title: "simple-usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simple-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

mar_par <- par("mar")
```

# Introduction

The `HurstDCWT` package can be used to estimate the Hurst exponent of an image using discrete complex wavelet transforms (DCWT). In this vignette, we will demonstrate the basic usage in a couple examples.

First, load the relevant packages, and define a quick function that we'll use for displaying the grayscale images that we will process later.

```{r setup}
library(HurstDCWT)

plot_square_image <- function(image_mat) {
  par(mar = c(0, 0, 0, 0))
  image(t(image_mat[nrow(image_mat):1, ]), 
        col = gray.colors(256), asp = 1, axes = FALSE)
}
```

# A boat!

First, we'll look at an example image and see what happens when we process it with the DCWT. This image of a boat is from the [USC Signal and Image Processing Institute image database](https://sipi.usc.edu/database/).

```{r, out.width="67%"}
data(boat)
plot_square_image(boat)
```


```{r, out.width="67%"}
boat_dwt_ex <- image_dwt_mult(boat, 2, 
                              family = "LinaMayrand", filter.number = 5.1)
boat_dwt_plot <- Mod(boat_dwt_ex$dwt)
plot_square_image(boat_dwt_plot)
```

```{r}
boat_dwt <- image_dwt_mult(boat, 8, 
                              family = "LinaMayrand", filter.number = 5.1)
# use the "access" function to get one of the regions
# roughly give a sense of the "size" of coefficients
summary(as.vector(Mod(access_dwt2D(boat_dwt, 1, "diagonal"))))
```

```{r}
data(fbm2d_512_75)

# an example of fractional Brownian motion with H = 0.75
plot_square_image(fbm2d_512_75)
```
```{r}
fbm_dwt <- image_dwt_mult(fbm2d_512_75, 8, 
                          family = "LinaMayrand", filter.number = 5.1)

fbm_energies <- get_energies(fbm_dwt, location_stat = "mean")
fbm_slope_params <- get_slope(fbm_energies)

plot_spectra(fbm_energies)
```

```{r}
est_H(fbm_slope_params)
```

```{r end, include=FALSE}
par(mar = mar_par)
```

